Laves polyhedrons in synthetic tennantite, Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}, and the lattice dynamics

### 2021-01-24 paper_tennantite.tex

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

@@

The tennantite-tetrahedride group represents a large class of inorganic compounds including complex copper sulfosalts with the general formula Cu\textsubscript{10+x}M\textsubscript{2-x}X\textsubscript{4}S\textsubscript{13}, where M is a transition metal (Zn, Fe, Hg, etc.), and X is a pnictogen (As, Sb, or Bi) \cite{Makovicky_2006}, with the pure tennantite itself represented by the formula Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}.

Despite being known for over two centuries, these minerals are still under investigation, both due to  structural peculiarities and potential applications. Arsenic- and antimony-containing compounds of this class are the advanced thermoelectric materials with a low lattice thermal conductivity\cite{Sootsman2009,Chetty2015}. Understanding the relationships between the crystal structure and phonon propagation, and scattering, is of a prime importance for designing new thermoelectric materials.

In a case of tennantite- and tetrahedrite-based materials, it is generally considered, that a low lattice thermal conductivity is caused by hybridized Cu {\it 3d}  and chalcogen p orbitals\cite{Lai2015}.
Such a hybridization leads to a bonding asymmetry, and out-of-plane low frequency rattling modes, which are quasi-localized and anharmonic\cite{May2016,Bouyrie2015}.

>>>>

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

@@

The tennantite-tetrahedride group represents a large class of inorganic compounds including complex copper sulfosalts with the general formula Cu\textsubscript{10+x}M\textsubscript{2-x}X\textsubscript{4}S\textsubscript{13}, where M is a transition metal (Zn, Fe, Hg, etc.), and X is a pnictogen (As, Sb, or Bi) \cite{Makovicky_2006}, with the pure tennantite itself represented by the formula Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}.
Despite being known for over two centuries, these minerals are still under investigation, both due to  structural peculiarities and potential applications. Arsenic- and antimony-containing compounds of this class are the advanced thermoelectric materials with a low lattice thermal conductivity\cite{Sootsman2009,Chetty2015}.

\textcolor{red}{The tennantite structure treated as sphalerite-type and belongs to \hmn{I -4 3 m} space group\cite{yaroslavzev2019,Makovicky_2006}.
There are three copper position: Cu1 is tetrahedrally coordinated by four sulpfur atoms, Cu2 is coordinated by three sulpfur atoms and Cu21 is tetrahedrally coordinated by three sulpfur and arsenic atoms.
Cu21 and Cu2 atoms possess very large atomic displacements perpendicular to the plane of the triangle (Fig.~\ref{fig:generel_view} right).}
In a case of tennantite- and tetrahedrite-based materials, it is generally considered, that a low lattice thermal conductivity is caused by hybridized Cu {\it 3d}  and chalcogen p orbitals\cite{Lai2015}.
Such a hybridization leads to a bonding asymmetry, and out-of-plane low frequency rattling modes, which are quasi-localized and anharmonic\cite{May2016,Bouyrie2015}.

\begin{figure}[ht]
\centering
\subfigure{\includegraphics[width=0.4\columnwidth]{general_view}}
 \quad
\subfigure{\includegraphics[width=0.35\columnwidth]{general_view_laves}}
\caption{\label{fig:generel_view} \textcolor{red}{Left --- the view of the Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} tennantite structure and of coordination polyhedra around the Cu1 (blue) and the Laves polyhedra (yellow); right --- the Laves polyhedra with Cu2 and Cu21 atoms with occupancy.} }
\end{figure}
